Spin canonical divisors of log stacky curves

### 2015-06-27 stacky_write_up/spin_cring.tex
comments to page 3 of induction section

As suggested by the previous lemma, \todo{Reference the lemma, instead of using previous, since things might get moved around} we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma. \todo{Again, reference this}
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, \todo{attached to $x_i$ doesn't really sound like the correct terminology to me} the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} \todo{This name seems a bit funny looking to me} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff \todo{It's better not to use abbreviations like iff in papers. Also, usually in definitions, one uses if in place of iff.} one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$ \todo{Order at which point, in this line and the next}
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative. \todo{I do not understand what this means, maximal under what ordering? Also, }
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is, Aaron: I suggest not to use the phrase ``leave as an exercise'' in a paper}
We now offer a lemma that allows us to inductively add $\mathbb{Q}$-points to a divisor and describe the new generators and relations.

\begin{lem}
\label{lem:sat-1}
Let $X$ be a genus $g$ curve with a $\BQ$-divisor $\halfcan'$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generically chosen generators (in the necessary degrees)
$u, x_1, \ldots, x_m$ with $\ord_P(x_i)=0$ for all $x_i$, $\ord_P(u) = 0$, \todo{Aaron: Instead of having $\deg u = 0$, don't we only need that $\ord_P(u) = \lfloor \halfcan'|_P \rfloor$?} and $\deg(u)= 1$ when viewed as sections of $\so_{\halfcan'}$.  Suppose $u, x_1, \ldots, x_m$ generate $R_{\halfcan'}$ in degree at most $\sigma$ with relations generated
in degree at most $\tau$.  Suppose $\halfcan = \halfcan' + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{d\halfcan}\rfloor) = h^0(X,\lfloor d\halfcan'
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then

\begin{enumerate}
\item[(a)] $R_{\halfcan}$ is generated over $R_{\halfcan'}$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$, when viewed as sections of $\so_{\halfcan'}$

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
\[
	\ord_u(f) < \ord_u(g) \implies f\prec g
\]
and equip $k[y_1, \ldots, y_n]$ with g-P-revlex and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ with a block order.
If $I$ and $I'$ are the ideal of relations of $R_{\halfcan}$ and $R_{\halfcan'}$ respectively, then

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle U_i: 1 \le i \le n-1 \rangle\\
										 &+ \langle V \rangle\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_i^{a_i}$ with $a_j\in \mathbb{Z}^{\ge 0}, a_i\ne 0$ such that
$\sum_{j=1}^i a_j c_j \ge c_{j+1}$
and there does not exist $((b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ such that with all $b_j\le a_j$ and 
$\sum_{j=1}^i b_jc_j\ge c_{j+1})\}$.  \\

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with relations $I$ generated over $I'$ by relations in degree is generated $\max(2\beta, \beta + \sigma)$ (where $\sigma=\max(\deg(x_i))$).
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
%\item[(c)] A Gr\"{o}bner basis with the initial terms $y_i x_j$, $y_i y_j$,
%	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
%	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
%generates $I'$.
\end{enumerate}
\end{lem}
\begin{proof}
By
our hypothesis (Equation ~\ref{eqn:deg1_sat_ind_dim}), for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

\noindent
so in $H^0 (X, \lfloor d\halfcan \rfloor)$ we can find elements with pole
order $i$ at $P$ for any $i \in \{0, \ldots, \lfloor d \frac{\alpha}{
\beta} \rfloor \}$. We will use O'Dorney's method of best lower approximations to
construct the $y_i$'s as desired. \todo{cite where}

@@

\noindent
be the positive best lower approximation of $\frac{
\alpha}{\beta}$ as described in the lemma's statement.  Note that elements of $H^0(X, d\halfcan)$ have pole of
orders ranging from 0 to $\lfloor d_i \frac{\alpha}{\beta}\rfloor =
c_i $ at $P$.  Moreover, a set $\lfloor d_i \frac{\alpha}{\beta}\rfloor$ elements in $(R_{D})_{d_i}$ with respective pole orders $1, \ldots, \lfloor
d_i \frac{\alpha}{\beta} \rfloor$ are linearly independent over $(R_{\halfcan'})_d$, and thus by dimension counting form a $k$-basis for
$(R_{\halfcan})_d$ over $(R_{\halfcan'})_d$.

For any $j < i \in \mathbf{N}$, $\frac{c_i}{d_i} > \frac{c_j}{d_j}$, so for any $z_1, \ldots, z_n$ with $\deg(z_j)<d_i$ for all $z_j$, $\prod_{j=1}^n z_j$ must satisfy
\[
	\frac{-\ord_P(z_i)}{\deg(z_i)} < \frac{c_i}{d _i},
\]

\noindent
so we have the pole order of $\prod_{j=1}^n z_j$ is 
\[
	\sum -\ord(z_j) < c_i .
\]

@@

One key difference between the notion of admissibility in Definition ~\ref{defn:admissible} and \cite[Definition 8.5.1]{vzb:stacky} is that in \cite{vzb:stacky}, Voight and Zureick-Brown may assume that $\{P_i\} \cap \Supp(K_X) = \emptyset.$ However, we cannot assume that $\{P_i\} \cap \Supp(L_X) = \emptyset$, as $L_X$ is not necessarily basepoint-free. Therefore, we will have to work with orders of zeros of poles relative to the divisor we are inducting from, instead of relative to the structure sheaf. Hence, we introduce the following notation.

>>>>

As suggested by the previous lemma, \todo{Reference the lemma, instead of using previous, since things might get moved around} we will need an ordering on monomials, and the following definition will prove useful in the ensuing lemma. \todo{Again, reference this}
\begin{defn}
We equip the graded ring $k[x_1, \ldots x_m]$, with weight $a_i$ and pole order $-\ord_P(x_i)$ attached to $x_i$, \todo{attached to $x_i$ doesn't really sound like the correct terminology to me} the graded $P$ reverse lexicographic order {\bf{g-P-revlex}} \todo{This name seems a bit funny looking to me} if for $f=\prod_{i=1}^m {x_i}^{q_i}\ne g=\prod_{i=1}^m {x_i}^{r_i}$, $f\prec g$ iff \todo{It's better not to use abbreviations like iff in papers. Also, usually in definitions, one uses if in place of iff.} one of the following holds
\begin{enumerate}
\item $\deg(f) < \deg(g)$
\item $\deg(f) = \deg(g)$ and $-\ord(f) > -\ord(g)$ \todo{Order at which point, in this line and the next}
\item $\deg(f) = \deg(g)$, $-\ord(f) = -\ord(g)$, and for $i$ maximal such that $q_i-r_i$ is non-zero, that expression is negative. \todo{I do not understand what this means, maximal under what ordering? Also, what expression are you talking about?}
\end{enumerate}
\end{defn}
We leave it as an exercise to verify that this is an order.  \todo{check for ourselves that this actually is, Aaron: I suggest not to use the phrase ``leave as an exercise'' in a paper.}
We now offer \todo{Maybe state instead of offer?} a lemma that allows us to inductively add $\mathbb{Q}$-points \todo{$\BQ$ points usually refers to maps to $\spec \BQ$. Maybe we can just say a rational multiple of a poitn?} to a divisor and describe the new generators and relations.

\begin{lem}
\label{lem:sat-1}
Let $X$ be a genus $g$ curve and $\halfcan' \in \di X \otimes \BQ$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generators (in the necessary degrees \todo{What does this mean?})
$u, x_1, \ldots, x_m$ with $\deg u = 1$, $\ord_P^{\halfcan'}(x_i)=0$ for all $1 \leq i \leq m$, and $\ord_P^{\halfcan'}(u) = 0$.Suppose $u, x_1, \ldots, x_m$ generate $R_{\halfcan'}$ in degree at most $\sigma$ with relations generated
in degree at most $\tau$.  Suppose $\halfcan = \halfcan' + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{d\halfcan}\rfloor) = h^0(X,\lfloor d\halfcan'
	\rfloor) + \left\lfloor d\frac{\alpha} {\beta} \right \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then,

\begin{enumerate}
\item[(a)] $R_{\halfcan}$ is generated over $R_{\halfcan'}$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$, when viewed as sections of $\so_{\halfcan'}$ \todo{Clearer to say $\deg (y_i) = d_i,\ord_P^{L'} = c_i.$}

\item[(b)] Choose an ordering $\prec$ on $k[u, x_1, \ldots, x_m]$ such that
\[
	\ord_u(f) < \ord_u(g) \implies f\prec g
\]
and equip $k[y_1, \ldots, y_n]$ with g-P-revlex and $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ with a block order \todo{it should be called block order, instead of ``a block order''}.
If $I$ and $I'$ are the ideal of relations of $R_{\halfcan}$ and $R_{\halfcan'}$ respectively, then

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_i^{a_i}$ \todo{should this be $y_j^{a_j}$ instead of $y_i^{a_i}$?} with $a_j\in \mathbb{Z}^{\ge 0}, a_i\ne 0$ \todo{this conflation of $j$ and $i$ is confusing me, is $j$ ranging over all indices from $1$ to $i$?} such that
$\sum_{j=1}^i a_j c_j \ge c_{j+1}$
and there does not exist $((b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ such that with all $b_j\le a_j$ and 
$\sum_{j=1}^i b_jc_j\ge c_{j+1})\}$. \todo{what is this close bracket closing?}  \\

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with relations $I$ generated over $I'$ by relations in degree is generated $\max(2\beta, \beta + \sigma)$ {\rm(}where $\sigma=\max(\deg(x_i))$ \todo{include what indexing set this $i$ is over ($ i = 1,\ldots, m$).}{\rm)}.
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
%\item[(c)] A Gr\"{o}bner basis with the initial terms $y_i x_j$, $y_i y_j$,
%	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
%	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
%generates $I'$.
\end{enumerate}
\end{lem}
\begin{proof}
By
Equation ~\ref{eqn:deg1_sat_ind_dim}, for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

\noindent
\todo{why is there no floor in the above equation, as in ~\ref{eqn:deg1_sat_ind_dim}} so in $H^0 (X, \lfloor d\halfcan \rfloor)$ we can find elements with pole
order $i$ at $P$ for any $i \in \{0, \ldots, \lfloor d \frac{\alpha}{
\beta} \rfloor \}$. We will use O'Dorney's method of best lower approximations to
construct the $y_i$'s as desired. \todo{cite where} \todo{What do we desire about the $y_i$?}

@@

\noindent
be the positive best lower approximations of $\frac{
\alpha}{\beta}$ as described in this lemma's statement. Note that elements of $H^0(X, d\halfcan)$ have poles of
orders ranging from 0 to $\lfloor d_i \frac{\alpha}{\beta}\rfloor =
c_i $ at $P$.  Moreover, a set $\lfloor d_i \frac{\alpha}{\beta}\rfloor$ elements in $(R_{D})_{d_i}$ with respective pole orders $1, \ldots, \lfloor
d_i \frac{\alpha}{\beta} \rfloor$ are linearly independent over $(R_{\halfcan'})_d$, and thus by dimension counting form a $k$-basis for
$(R_{\halfcan})_d$ over $(R_{\halfcan'})_d$.

For any $j < i \in \mathbf{N}$, $\frac{c_i}{d_i} > \frac{c_j}{d_j}$, so for any $z_1, \ldots, z_n$ with $\deg(z_j)<d_i$ for all $z_j$, $\prod_{j=1}^n z_j$ must satisfy
\todo{In the next equations, are we viewing these as rational functions on the structure sheaf or on $L'$, what point is the order of $z_j$ taken with respect to?}
\[
	\frac{-\ord_P(z_i)}{\deg(z_i)} < \frac{c_i}{d_i},
\]

\noindent
so we have the pole order of $\prod_{j=1}^n z_j$ is \todo{what is the following sum taken over?}
\[
	\sum -\ord(z_j) < c_i .
\]

@@

One key difference between the notion of admissibility in Definition ~\ref{defn:admissible} and \cite[Definition 8.5.1]{vzb:stacky} is that in \cite{vzb:stacky}, Voight and Zureick-Brown may assume that $\{P_i\} \cap \Supp(K_X) = \emptyset.$ However, we cannot assume that $\{P_i\} \cap \Supp(L_X) = \emptyset$, as $L_X$ is not necessarily basepoint-free. Therefore, we will have to work with orders of zeros of poles relative to the divisor we are inducting from, instead of relative to the structure sheaf.
